\documentclass[11pt]{article}

\usepackage{times}


%% Page layout
\oddsidemargin 0pt
\evensidemargin 0pt
\textheight 600pt
\textwidth 469pt
\setlength{\parindent}{0em}
\setlength{\parskip}{1ex}

\begin{document}

\title{15-213, Fall 20xx\\
Data Lab: Manipulating Bits\\
Assigned: Aug.~30, Due: Wed., Sept.~12, 11:59PM
}

\author{}
\date{}

\maketitle

Harry Bovik ({\tt bovik@cs.cmu.edu}) is the lead person for
this assignment.

\section{Introduction}
The purpose of this assignment is to become more familiar with
bit-level representations of integers and floating point numbers.
You'll do this by solving a series of programming ``puzzles.'' Many of
these puzzles are quite artificial, but you'll find yourself thinking
much more about bits in working your way through them.

\section{Logistics}

This is an individual project. All handins are electronic.
Clarifications and corrections will be posted on the course Web page.

\section{Handout Instructions}

\begin{quote}
\bf SITE-SPECIFIC: Insert a paragraph here that explains how the
instructor will hand out the {\tt datalab-handout.tar} file to the
students.
\end{quote}

Start by copying {\tt datalab-handout.tar} to a (protected) directory
on a Linux machine in which you plan to do your work.  Then give the
command
\begin{verbatim}
  unix> tar xvf datalab-handout.tar.  
\end{verbatim}
This will cause a number of files to be unpacked in the directory.
The only file you will be modifying and turning in is {\tt bits.c}.

The {\tt bits.c} file contains a skeleton for each of the 13
programming puzzles.  Your assignment is to complete each function
skeleton using only {\em straightline} code for the integer puzzles
(i.e., no loops or conditionals) and a limited number of C arithmetic
and logical operators. Specifically, you are {\em only} allowed to use
the following eight operators:
\begin{verbatim}
 ! ~ & ^ | + << >>
\end{verbatim}
A few of the functions further restrict this list.  Also, you are not
allowed to use any constants longer than 8 bits.  See the comments in
{\tt bits.c} for detailed rules and a discussion of the desired coding
style.

\section{The Puzzles}

This section describes the puzzles that you will be solving in {\tt bits.c}.

Table \ref{puzzles-tab} lists the puzzles in rought order of
difficulty from easiest to hardest. The ``Rating'' field gives the
difficulty rating (the number of points) for the puzzle, and the ``Max
ops'' field gives the maximum number of operators you are allowed to
use to implement each function.  See the comments in {\tt bits.c} for
more details on the desired behavior of the functions. You may also
refer to the test functions in {\tt tests.c}.  These are used as
reference functions to express the correct behavior of your functions,
although they don't satisfy the coding rules for your functions.

\begin{table}[htbp]
\begin{center}
\begin{tabular}{|l|l|c|c|}
\hline
Name & Description & Rating & Max ops\\
\hline
{\tt bitXor(x,y)} & \verb@x || y@ using only \verb@&@ and \verb@~@. & 1 & 14\\
{\tt tmin()} & Smallest two's complement integer & 1 & 4\\
{\tt isTmax(x)} & True only if \verb@x@ x is largest two's comp. integer. &1&10\\
{\tt allOddBits(x)} & True only if all odd-numbered bits in \verb@x@ set to 1.  &2&12\\
{\tt negate(x)} & Return \verb@-x@ with using \verb@-@ operator.  & 2 & 5\\
{\tt isAsciDigit(x)} & True if \verb|0x30| $\leq$ \verb@x@ $\leq$.  &3&15\\
{\tt conditional} & Same as \verb@ x ? y : z@  &3&16\\
{\tt isLessOrEqual(x, y)} & True if \verb@x@ $\leq$ \verb@y@, false otherwise &3&24\\
{\tt logicalNeg(x))} & Compute \verb@!x@ without using \verb@!@ operator. &4&12\\
{\tt howManyBits(x)} & Min. no. of bits to represent \verb@x@ in two's comp.  &4&90\\
{\tt floatScale2(uf)} & Return bit-level equiv. of \verb@2*f@ for f.p. arg. \verb@f@.  &4&30\\
{\tt floatFloat2Int(uf)} & Return bit-level equiv. of \verb@(int)f@ for f.p. arg. \verb@f@.  &4&30\\
{\tt floatPower2(x)} & Return bit-level equiv. of \verb@2.0^x@ for integer \verb@x@.  &4&30\\
\hline
\end{tabular}
\end{center}
\caption{Datalab puzzles. For the floating point puzzles, value
  \texttt{f} is the floating-point number having the same bit
  representation as the unsigned integer \texttt{uf}.}  
\label{puzzles-tab}
\end{table}

For the floating-point puzzles, you will implement some
common single-precision floating-point operations.  For these puzzles,
you are allowed to use standard control structures (conditionals,
loops), and you may use both {\tt int} and {\tt unsigned} data types,
including arbitrary unsigned and integer constants.  You may not use
any unions, structs, or arrays.  Most significantly, you may not use
any floating point data types, operations, or constants.  Instead, any
floating-point operand will be passed to the function as having type
{\tt unsigned}, and any returned floating-point value will be of type
{\tt unsigned}.  Your code should perform the bit manipulations that
implement the specified floating point operations.

The included program \texttt{fshow} helps you understand the structure
of floating point numbers. To compile \texttt{fshow}, switch to the
handout directory and type:
\begin{verbatim} 
  unix> make 
\end{verbatim}
You can use \texttt{fshow} to see what an arbitrary pattern 
represents as a floating-point number:
\begin{verbatim}
  unix> ./fshow 2080374784

  Floating point value 2.658455992e+36
  Bit Representation 0x7c000000, sign = 0, exponent = f8, fraction = 000000
  Normalized.  1.0000000000 X 2^(121)
\end{verbatim}
You can also give \texttt{fshow} hexadecimal and floating point
values, and it will decipher their bit structure.

\section{Evaluation}

Your score will be computed out of a maximum of 67 points based on the
following distribution:
\begin{description}
\item[36] Correctness points.
\item[26] Performance points.
\item[5] Style points.
\end{description}


{\em Correctness points.} The puzzles you must solve have been given a
difficulty rating between 1 and 4, such that their weighted sum totals
to 36.  We will evaluate your functions using the \texttt{btest}
program, which is described in the next section.  You will get full
credit for a puzzle if it passes all of the tests performed by
\texttt{btest}, and no credit otherwise.

{\em Performance points.}  Our main concern at this point in the course is
that you can get the right answer.  However, we want to instill in you
a sense of keeping things as short and simple as you can.
Furthermore, some of the puzzles can be solved by brute force, but we
want you to be more clever.  Thus, for each function we've established
a maximum number of operators that you are allowed to use for each
function. This limit is very generous and is designed only to catch
egregiously inefficient solutions.  You will receive two points for
each correct function that satisfies the operator limit.

{\em Style points.} Finally, we've reserved 5 points for a subjective
evaluation of the style of your solutions and your commenting.  Your
solutions should be as clean and straightforward as possible.  Your
comments should be informative, but they need not be extensive.

\section*{Autograding your work}

We have included some autograding tools in the handout directory ---
\texttt{btest}, \texttt{dlc}, and \texttt{driver.pl} --- to help you
check the correctness of your work.

\begin{itemize}
\item {\bf \texttt{btest:}} This program checks the functional correctness of
  the functions in {\tt bits.c}. To build and use it, type the
  following two commands:
\begin{verbatim}
  unix> make
  unix> ./btest
\end{verbatim}
Notice that you must rebuild \texttt{btest} each time you modify your {\tt
  bits.c} file. 

You'll find it helpful to work through the functions one at a time,
testing each one as you go.  You can use the {\tt -f} flag to instruct
\texttt{btest} to test only a single function:
\begin{verbatim}
  unix> ./btest -f bitXor
\end{verbatim}
You can feed it specific function arguments
using the option flags {\tt -1}, {\tt -2}, and {\tt -3}:
\begin{verbatim}
  unix> ./btest -f bitXor -1 4 -2 5
\end{verbatim}
Check the file {\tt README} for documentation on running the {\texttt
  btest} program.


\item {\bf \texttt{dlc}:} This is a modified version of an ANSI C compiler from
the MIT CILK group that you can use to check for compliance with the
coding rules for each puzzle. The typical usage is:
\begin{verbatim}
  unix> ./dlc bits.c
\end{verbatim}
The program runs silently unless it detects a problem, such as an
illegal operator, too many operators, or non-straightline code in the
integer puzzles.  Running with the \texttt{-e} switch:
\begin{verbatim}
  unix> ./dlc -e bits.c  
\end{verbatim}
causes \texttt{dlc} to print counts of the number of operators used by
each function. Type {\tt ./dlc -help} for a list of command line options. 

\item {\bf \texttt{driver.pl}:} This is a driver program that uses \texttt{btest}
and \texttt{dlc} to compute the correctness and performance points for
your solution. It takes no arguments:
\begin{verbatim}
  unix> ./driver.pl
\end{verbatim}
Your instructors will use \texttt{driver.pl} to evaluate your
solution.

\end{itemize}

\section{Handin Instructions}

\begin{quote}
\bf SITE-SPECIFIC: Insert text here that tells each student how to
hand in their {\tt bits.c} solution file at your school.
\end{quote}

\section{Advice}

\begin{itemize}

\item Don't include the \verb:<stdio.h>: header file in your {\tt
bits.c} file, as it confuses \texttt{dlc} and results in some
non-intuitive error messages. You will still be able to use {\tt
printf} in your {\tt bits.c} file for debugging without including the
\verb:<stdio.h>: header, although \texttt{gcc} will print a warning that
you can ignore.

\item The \texttt{dlc} program enforces a stricter form of C declarations
than is the case for C++ or that is enforced by \texttt{gcc}.  In
particular, any declaration must appear in a block (what you enclose
in curly braces) before any statement that is not a declaration.  For
example, it will complain about the following code:
\begin{verbatim}
  int foo(int x)
  {
    int a = x;
    a *= 3;     /* Statement that is not a declaration */
    int b = a;  /* ERROR: Declaration not allowed here */
  }
\end{verbatim}

\end{itemize}

\section{The ``Beat the Prof'' Contest}

For fun, we're offering an optional ``Beat the Prof'' contest that
allows you to compete with other students and the instructor to
develop the most efficient puzzles. The goal is to solve each Data Lab
puzzle using the fewest number of operators. Students who match or
beat the instructor's operator count for each puzzle are winners!

To submit your entry to the contest, type:
\begin{verbatim}
    unix> ./driver.pl -u ``Your Nickname''
\end{verbatim}
Nicknames are limited to 35 characters and can contain alphanumerics,
apostrophes, commas, periods, dashes, underscores, and ampersands.
You can submit as often as you like. Your most recent submission will
appear on a real-time scoreboard, identified only by your nickname.
You can view the scoreboard by pointing your browser at
\begin{verbatim}
    http://$SERVER_NAME:$REQUESTD_PORT
\end{verbatim}

\begin{quote}
\bf SITE-SPECIFIC: Replace \verb:$SERVER_NAME: and \verb:$REQUESTD_PORT:
with the values you set in the \verb:./contest/Contest.pm: file.
\end{quote}


\end{document}
